---
title: "Challenge 1"
author: "Pablo Barbera, Dan Cervone"
answers written by: "Miyabi Ishihara"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `congress-terms.csv`. Try not to look at the previous code we used!

```{r}
setwd("~/Desktop/Short Course/data-science-workshop-master/modeling-viz")
congress <- read.csv("./congress-terms.csv", stringsAsFactors = FALSE)
```


1. Make a time series plot showing Democratic/Republican control of congress over time. By varying color, line thickness, or other graphical properties, make this plot also convey the average age of the congressmen representing each part over time.

```{r}

num.dem <- tapply(congress$party, congress$congress, function(x) sum(x == "D"))
num.rep <- tapply(congress$party, congress$congress, function(x) sum(x == "R"))
names(num.dem)
num.congress <- as.numeric(names(num.dem))
num.congress

dems <- subset(congress, party == "D")
reps <- subset(congress, party == "R")

dems$age <- as.numeric(as.vector(dems$age))
reps$age <- as.numeric(as.vector(reps$age))

# calculating avg age of D and R for each congress
age.dem <- tapply(dems$age, dems$congress, function(x) mean(x, na.rem = T))
age.rep <- tapply(reps$age, reps$congress, function(x) mean(x, na.rem = T))

# creating a dataframe with necessary info
congress2 <- data.frame(num.congress, num.dem, num.rep, age.dem, age.rep)

# plots
plot(congress2$num.congress, congress2$num.dem - congress2$num.rep)
congress2 <- subset(congress2, num.congress >= 80)
plot(congress2$num.congress, congress2$num.dem - congress2$num.rep)

qplot(num.congress, num.dem - num.rep, data = congress2, color = age.dem, size = age.rep)

```

2. Write a function that, for a particular congress (e.g. congress 113), shows a map of the party breakdown by state.

```{r}
# creating a subset of congress 113 
congress$congress <- as.numeric(congress$congress)
class(congress$congress)
congress113 <- subset(congress, congress == 113)


# creating dataframe with the percentage of democrats for each state
num.dem.113 <- tapply(congress113$party, congress113$state, function(x) sum(x == "D"))
num.rep.113 <- tapply(congress113$party, congress113$state, function(x) sum(x == "R"))
  # d <- as.numeric(as.vector(num.dem.113)) checking if numbers add up
  # sum(d, na.rm = TRUE)
percent.dem <- num.dem.113/(num.dem.113 + num.rep.113)
state <- names(num.dem.113)
data <- data.frame(state, control = percent.dem)
data <- data[-c(1,2,3),]  # removing irrelevant rows


# getting state lat/lon data
library(maps)
states <- map_data("state") # long/lat information for all states
states$state <- state.abb[match(states$region, tolower(state.name))] # converting to abb


# merging data and mapping
states.data <- merge(states, data, by="state")
state.map <- 
    ggplot(states.data, aes(x=long, y=lat, group=group, fill=control)) + 
    geom_polygon(color="white", size = 0.2) + 
    theme_bw() + 
    scale_fill_gradient(low="red", high="blue", na.value="black") + 
    guides(fill=guide_legend(title="% Democratic")) +
  　coord_map("mercator") + 
    ggtitle(paste("% Democratic in Congress", 113, "by State"))
state.map 


# generalizing codes so that we get a map for any congress
library(maps)
states <- map_data("state") # long/lat information for all states
states$state <- state.abb[match(states$region, tolower(state.name))] # converting to abb

map <- function(cong){ # input congress number
  sub.congress <- subset(congress, congress == cong)
  
  num.dems <- tapply(sub.congress$party, sub.congress$state, function(x) sum(x == "D"))
  num.reps <- tapply(sub.congress$party, sub.congress$state, function(x) sum(x == "R"))
  percent.dems <- num.dems / (num.dems + num.reps)
  state <- names(num.dems)
  data <- data.frame(state, control = percent.dems)
  
  states.data <- merge(states, data, by="state")
  state.map <- 
    ggplot(states.data, aes(x=long, y=lat, group=group, fill=control)) + 
    geom_polygon(color="white", size = 0.2) + 
    theme_bw() + 
    scale_fill_gradient(low="red", high="blue", na.value="black") + 
    guides(fill=guide_legend(title="% Democratic")) +
  　coord_map("mercator") +
    ggtitle(paste("% Democratic in Congress", cong, "by State"))
  state.map 
}

map(113)

```