---
title: "Challenge 1"data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAAAkAAAAJCAYAAADgkQYQAAAAMElEQVR42mNgIAXY2Nj8x8cHC8AwMl9XVxe3QqwKcJmIVwFWhehW4LQSXQCnm3ABAHD6MDrmRgfrAAAAAElFTkSuQmCC
author: "Pablo Barbera, Dan Cervone"
answers written by: "Miyabi Ishihara"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `humansofnewyork.csv`. Read this data into `R` and answer the questions below, trying not to rely too much on the previous code we used!

```{r}
# set working directory before reading data
humans <- read.csv("./humansofnewyork.csv", stringsAsFactors = FALSE) 
str(humans)
dim(humans) 
head(humans) 
summary(humans)
```


1. How many status updates have been posted on this page?

```{r}
table(humans$type)
sum(humans$type == "status") # 83
```

2. What is the total number of likes, comments, and shares it received?

```{r}
sum(humans$likes_count)
sum(humans$comments_count)
sum(humans$shares_count, na.rm = TRUE)

```

3. What is the content of the post with the highest number of shares?

```{r}
maxshare = max(humans$shares_count, na.rm = TRUE) 
humans$message[which(humans$shares_count == maxshare)] 


# another way is by sorting
sorted = sort(humans$shares_count, decreasing = TRUE)
sorted[1]
humans$message[which(humans$shares_count == sorted[1])]

```

4. What was the date in which the first photo was posted?

```{r}
photos <- subset(humans, humans$type == "photo")
sorted.photos <- sort(t$created_time, decreasing = FALSE)
sorted.photos[1] # October 11, 2011


# another way
first.photo <- min(which(humans$type == "photo")) # storing id of the first photo
humans$created_time[first.photo]


# unnecessarily long way
type.table <- table(humans$type)
type.names <- names(type.table)
type.names[grep("photo", type.names)]
photos <- subset(humans, type %in% type.names[grep("photo", type.names)])
sorted = sort(photos$created_time, decreasing = FALSE)
sorted[1]

```

5. What is the total number of likes that the page has ever received, excluding its most popular post?

```{r}
total.likes <- sum(humans$likes_count)
max.likes <- max(humans$likes_count) # no. of likes of a most popular post
total.likes - max.likes
```

6. How many posts have received more than 1,000,000 likes?

```{r}
sum(humans$likes_count > 1000000)
```

7. What was the total number of shares received by posts published each year?


```{r}

humans$year <- substr(humans$created_time, 1, 4)
tapply(humans$shares_count, humans$year, sum)  # tapply is very useful

```
