---
title: "Challenge 2"
author: "Pablo Barbera, Dan Cervone"
edited by: "Miyabi Ishihara"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `congress-terms.csv`. Please read the data into `R` and answer the questions below.

```{r}
# set working directory before reading data
congress <- read.csv("./congress-terms.csv", stringsAsFactors = FALSE)

```


1. How many total congressmen have represented each political party?

```{r}
table(congress$party)
congress <- subset(congress, party %in% c("D", "I", "R")) # keeping only dem, ind, rep
table(congress$party)

```

2. Who was the youngest congressman elected and in how many congresses did he/she serve?
ANSWER: Jed Johnson, 25 years old, 1 congress

```{r}
min(congress$age)
youngest <- which.min(congress$age)
first.youngest <- congress$firstname[youngest]
last.youngest <- congress$lastname[youngest]
paste(first.youngest, last.youngest)
sum(congress$firstname == first.youngest & congress$lastname == last.youngest)


# another way
age <- as.numeric(as.vector(congress$age))
congress$age <- age
minage = min(congress$age, na.rm=TRUE)
minage 

firstname = (congress$firstname[which(congress$age == minage)])
lastname = (congress$lastname[which(congress$age == minage)])
```


3. In how many congresses did the House of Representatives have a Democratic majority?

```{r}

house <- subset(congress, chamber == "house")
house.dems <- tapply(house$party, house$congress, function(x) sum(x == "D"))
house.reps <- tapply(house$party, house$congress, function(x) sum(x == "R"))
sum(house.dems > house.reps)

```

4. Which state has had the oldest congressmen on average?

```{r}

congress$age <- as.numeric(congress$age)
age.table <- tapply(congress$age, congress$state, mean)
oldest <- sort(age.table, decreasing = TRUE)[1]
age.table[which.max(age.table)] # another way

```

5. For which state is the proportion of total Democrats throughout history closest to the proportion of Republicans?

```{r}

dem <- tapply(congress$party, congress$state, function(x) sum(x == "D"))
rep <- tapply(congress$party, congress$state, function(x) sum(x == "R"))
prop.dem <- dem/(dem + rep)
diff <- abs(prop.dem - 0.5)
diff[which.min(diff)]

sort(diff, decreasing = FALSE)[1] # another way to do this is by sorting

```

6. How many times has the majority party in the House of Representatives changed?

```{r}

dem.majority <- house.dems > house.reps
party.change <- 0
for(i in 2:length(dem.majority)){
  if(dem.majority[i] != dem.majority[i-1]){
    party.change <- party.change + 1
  }
}
party.change #6
```

7. Which state in the 113th congress has the most experienced congressmen (measured my the average number of congresses in which they served)?

```{r}

experience <- rep(0, nrow(congress))
for(i in 2:nrow(congress)){
  # counting the no. of previous appearances of each congressman
  experience[i] <- sum(congress$firstname[1:(i-1)] == congress$firstname[i] & congress$lastname[1:(i-1)] == congress$lastname[i])
}

congress$experience <- experience
congress.113 <- subset(congress, congress == 113)
# mean experience year for each state in congress.113
exp.table <- tapply(congress.113$experience, congress.113$state, mean)
exp.table[which.max(exp.table)]

# ans: Vermont (average number served is 11 years)
```
