---
title: "Challenge 1"
author: Miyabi Ishihara
output: html_document
---

Problems created by Pablo Barbera, Dan Cervone
[&laquo; Text Analysis Module](../text/README.md)

Write commands that help you answer the following questions about the bullying dataset.

1. How many of these tweets are in English? (there is a column that can help you here)

```{r, results = "hide"}
df.tweets <- read.csv("bullying.csv", header = TRUE, stringsAsFactors = FALSE)
table(df.tweets$lang)
table(df.tweets$lang)["en"]
```


2. How many of the tweets contain a link? How many mention another Twitter user?
```{r, results = "hide"}
length(grep("HTTPLINK", df.tweets$text))
length(grep("@USER", df.tweets$text))
```


3. How many tweets contain a hashtag?
```{r, results = "hide"}
length(grep("#", df.tweets$text))
```


4. How many of the tweets mention anything about a school? (include any word you think relate to school)
```{r, results = "hide"}
length(grep("school|class|teacher|student", df.tweets$text, ignore.case = TRUE))
```


5. Display the first 20 tweets which mention a school.
```{r, results = "hide"}
df.tweets$school <- grepl("school|class|teacher|student", df.tweets$text, ignore.case = TRUE)
school.tweets <- df.tweets[df.tweets$school == TRUE,]
school.tweets$text[1:20]
school.tweets[,c("text")][1:20] # another way to do the same thing
```


6. How many tweets mention both "mom" and "dad"?
```{r, results = "hide"}
df.tweets$mom <- grepl(" mom |^mom | mom| mom\\.", df.tweets$text, ignore.case=TRUE)
df.tweets$dad <- grepl(" dad |^dad | dad$| dad\\.", df.tweets$text, ignore.case=TRUE)

table(df.tweets$mom == TRUE & df.tweets$dad == TRUE)
df.tweets$text[df.tweets$mom & df.tweets$dad]
```


7. Display the first 15 tweets which mention "mom" and "dad". Note any peculiar findings.

```{r, results = "hide"}
head(df.tweets$text[df.tweets$mom==TRUE & df.tweets$dad==TRUE], n=15)
```

